---
title: "a7_networks_jenny"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(readxl)
library(tidyverse)
library(viridis)
library(patchwork)
library(hrbrthemes)
library(circlize)
library(chorddiag)  #devtools::install_github("mattflor/chorddiag")
```

You have to extract the data from this Excel file programmatically, i.e., using R and/or Python, not just copy it by hand to another Excel file. For this, you will have to open and understand the structure and format of the data in the file.
```{r}
migra = read_excel("./all_dataset/A7_data/State_to_State_Migrations_Table_2018.xls")
##data(state)
c = apply(migra[6,],1,function(x)trimws(x, "r")) #remove right-side spaces
col_ix = which(c %in% state.name)
r = apply(migra[,1],2,function(x)trimws(x, "r"))
row_ix = which(r %in% state.name)
df <- data.frame(migra[row_ix,col_ix])
rownames(df) = migra[row_ix,1][[1]]
colnames(df) = unlist(migra[6,col_ix])
##head(df)
```

You will need to extract the data in the form of a data frame with 3 columns: source, target and count; the last column will contain the numbers of people who migrated from the source state to the target state in 2018.
```{r}
# convert paired matrix to long format:https://www.data-to-viz.com/graph/chord.html
statename_df <- as.data.frame(state.name)
stateindex <- c(1:length(state.name))
rownames(statename_df) <- stateindex
###
network <- expand.grid(stateindex,stateindex) #all pairs of state names
colnames(network) <- c("source","target")
network$count <-NA
for (r in c(1:nrow(network))){
  s <- as.numeric(network$source[r])
  t <- as.numeric(network$target[r])
  if(length(df[s,t])>0){network$count[r] <- df[s,t]}
  network$source[r] <- statename_df[s,]
  network$target[r] <- statename_df[t,]
}
###
```

You can now use this data frame to create a graph object using igraph (R or Python), tidygraph (R) or networkx (Python).
You will create a single network graph, where the width of each edge representing the number of people moving between the two states, the color of the edge denoting the state of origin, and preferably arrows denoting the direction of migration. These aspects are required. You can style other aspects of the graph as you choose to improve readability.
```{r}
plot_state <- c("Texas","Arizona","Washington","Nevada","Oregon","Florida","Colorado","Utah","New York","Georgia","New Jersey","California","Pennsylvania","North Carolina","Connecticut","Massachusetts","Virginia")
network_plot <- network[network$source %in% plot_state & network$target %in% plot_state,]
####
# parameters
circos.clear()
circos.par(start.degree = 90, gap.degree = 4, track.margin = c(-0.1, 0.1), points.overflow.warning = FALSE)
par(mar = rep(0, 4))

# color palette
mycolor <- viridis(length(plot_state), alpha = 1, begin = 0, end = 1, option = "D")
mycolor <- mycolor[sample(1:length(plot_state))] #17 is the number of states

# Base plot
chordDiagram(
  x = network_plot, 
  grid.col = mycolor,
  transparency = 0.25,
  directional = 1,
  direction.type = c("arrows", "diffHeight"), 
  diffHeight  = -0.04,
  annotationTrack = "grid", 
  annotationTrackHeight = c(0.05, 0.1),
  link.arr.type = "big.arrow", 
  link.sort = TRUE, 
  link.largest.ontop = TRUE)

# Add text and axis
circos.trackPlotRegion(
  track.index = 1, 
  bg.border = NA, 
  panel.fun = function(x, y) {
    
    xlim = get.cell.meta.data("xlim")
    sector.index = get.cell.meta.data("sector.index")
    
    # Add names to the sector. 
    circos.text(
      x = mean(xlim), 
      y = 3.2, 
      labels = sector.index, 
      facing = "bending", 
      cex = 0.8
    )
  }
)
```

